---
title: "SwarmSVM"
output: 
  rmarkdown::html_vignette:
    number_sections: yes
    toc: yes
author: Tong He
vignette: >
  %\VignetteIndexEntry{SwarmSVM}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Introduction
=============

This package contains several ensemble learning algorithms based on the following papers:

1. Gu, Q., & Han, J. (2013). Clustered support vector machines. In proceedings of the sixteenth international conference on artificial intelligence and statistics (pp. 307-315).
2. Hsieh, C. J., Si, S., & Dhillon, I. S. (2013). A divide-and-conquer solver for kernel support vector machines. arXiv preprint arXiv:1311.0914.
3. Collobert, R., Bengio, S., & Bengio, Y. (2002). A parallel mixture of SVMs for very large scale problems. Neural computation, 14(5), 1105-1114.

The main idea of these algorithms are

1. Reducing the scale of the data set results in a faster algorithm.
2. If we divide a linear inseperable problem into smaller sub-problems appropriately, then it is possible to solve it by linear SVMs.

These two ideas focus on the efficiency and accuracy respectively. Specifically, we usually use more than one SVM model to solve the whole problem, therefore this is also an ensemble learning framework.


Data
==================================

In this package, we choose a small data set to demonstrate the usage of our functions. The data set is `svmguide1` from [libsvm's official website](http://www.csie.ntu.edu.tw/~cjlin/libsvmtools/datasets/binary.html). The data is collected from an astroparticle application from Jan Conrad of Uppsala University, Sweden.

We can load it by 

```{r, message=FALSE}
require(SwarmSVM)
data(svmguide1)
```

It is a list object. Let's first take a look at the training data.

```{r}
head(svmguide1[[1]])
```

The first column contains the classification target value, the other columns contain the features. It is a binary classification task. The second part in the list is the test set: 

```{r}
head(svmguide1[[2]])
```

We rename them with the following command:

```{r}
svmguide1.t = svmguide1[[2]]
svmguide1 = svmguide1[[1]]
```

From now on, we have the training data set `svmguide1` and the test data set `svmguide1.t`.

Clustering Algorithm
==================================

In our pacakge, there are two main algorithms requiring clustering algorithm for the input data. Therefore we provide some clustering algorithms for users to choose from. Users can also implement their own functions and pass it to our algorithm.

#### Default Algorithms

We now provide two algorithms existing in R:

- The default kmeans algorithm `stats::kmeans`, named as "kmeans";
- The kernel kmeans algorithm `kernlab::kkmeans`, named as "kernkmeans".

In `clusterSVM` and `dcSVM`, we offer an argument `cluster.method`, you could choose one of the two algorithms and pass its name to the argument.

#### Customization

We also offer arguments for users to pass their own implementation of the clustering algorithm: `cluster.fun` and `cluster.predict`.

`cluster.fun` is the clustering training function. It takes a function requiring the data and number of centers as the two main arguments. The output of this function should be an list object of the clustering result, while it has two fields:

1. `object$cluster` as the clustering label on the input data.
2. `object$centers` as the clustering center matrix.

`cluster.predict` is the predicting algorithm on the trained clustering object. It takes a function requiring the data and trained clustering object from `cluster.fun` as the two main arguments. The output of this function should be simply a vector of tue clustering label on the input data.


Clustered Support Vector Machines
==================================

#### Algorithm

The algorithm is straight forward:

**Training**

1. Cluster the data. The default setting is `stats::kmeans`.
2. Transform the data according to the Eq. (4) - Eq. (7) in the original paper.
3. Solve the new problem with a linear svm from `LiblineaR::LiblineaR`.

**Test**

1. Assign cluster label to each new data point, based on the clustering result from training.
2. Transform the data according to the Eq. (4) - Eq. (7) in the original paper.
3. Make prediction with the trained model.

#### Basic usage

We demonstrate the usage of this function with the following code:

```{r}
csvm.obj = clusterSVM(x = svmguide1[,-1], y = svmguide1[,1], type = 1,
                      valid.x = svmguide1.t[,-1],valid.y = svmguide1.t[,1], 
                      seed = 1, verbose = 1, centers = 8)
csvm.obj$valid.score
```

Here the parameters are grouped into four parts:

1. `x` and `y` are the feature matric and target vector of the training data. `type` is specifying the mission and the type of the SVM.
2. `valid.x` and `valid.y` are the feature matric and target vector of the validation data.
3. `seed` is controlling the random seed to make the result reproducible. `verbose` is controlling the content of the output.
4. `centers` is the parameter passing to the cluster algorithm.

**Dense and sparse input**

The sample data set is in the format of sparse matrix.

```{r}
class(svmguide1)
```

The function takes a dense matrix or a sparse matrix as the input feature matrix. Therefore the following code gives you the same result.

```{r}
csvm.obj = clusterSVM(x = as.matrix(svmguide1[,-1]), y = svmguide1[,1], type = 1,
                      valid.x = as.matrix(svmguide1.t[,-1]),valid.y = svmguide1.t[,1], 
                      seed = 1, verbose = 1, centers = 8)
csvm.obj$valid.score
```

**Self-defined clustering algorithm**

In `clusterSVM`, the clustering is a very important step. Therefore we don't restrict users to the `RcppMLPACK::mlKmeans` algorithm. Instead, we accept user-defined clustering algorithm as an argument.

Note that we require the output of the clustering algorithm contains two fields: `centers` and `cluster`. One example could be

```{r}
cluster.fun = stats::kmeans

cluster.predict = function(x, cluster.object) {
  centers = cluster.object$centers
  eucliDist = function(x, centers) apply(centers, 1, function(C) colSums( (t(x)-C)^2 ))
  euclidean.dist = eucliDist(x, centers)
  result = max.col(-euclidean.dist)
  return(result)
}
```

Here we use the default kmeans, and implement the prediction function. Once we have defined the algorithm, it is straight forward to pass it to `clusterSVM`:

```{r}
csvm.obj = clusterSVM(x = svmguide1[,-1], y = svmguide1[,1], centers = 8, seed = 1,
                      cluster.fun = cluster.fun, cluster.predict = cluster.predict,
                      valid.x = svmguide1.t[,-1], valid.y = svmguide1.t[,1])
csvm.obj$valid.score
```


A Divide-and-Conquer Support Vector Machine
=============================================

#### Algorithm

The algorithm could be described as the following:

**Training**

1. We cluster the data in a recursive manner: 
    - Cluster the data in `k` groups.
    - In each groups, we keep cluster the data into `k` finer groups.
    - Repeat for `max.levels` times
2. At the finest level, we train svm models
3. At the `j`-th group on level `l`, we
    - Concatenate the `alpha` (coefficient on support vector) value from the subgroups of the `j`-th group
    - Train the svm model with the `alpha` value initialized
4. Fine tune the `alpha` values by training an svm on all the support vectors of the whole data set.
5. Train the final svm model with the `alpha` value

**Test**

There are two ways to do prediction.

1. Early Prediction
    - In early predicton on level `l`, we predict the clustering label at level `l` for the new input data.
    - For data points belong to each subgroup, we make prediction using the corresponding svm model for this subgroup.
2. Exact Prediction
    - We predict with the final svm model on the new input data.

#### Basic usage

We demonstrate the usage of this function with the following code:

```{r}
dcsvm.model = dcSVM(x = svmguide1[,-1], y = svmguide1[,1],
                    k = 4, max.levels = 4, seed = 0, cost = 32, gamma = 2,
                    kernel = 3,early = 0, m = 800,
                    valid.x = svmguide1.t[,-1], valid.y = svmguide1.t[,1])
dcsvm.model$valid.score
```

Here the parameters can be grouped into five parts.

1. `x` and `y` are the feature matric and target vector of the training data. 
2. `valid.x` and `valid.y` are the feature matric and target vector of the validation data.
3. `seed` is controlling the random seed to make the result reproducible.
4. `k`, `max.levels` controls the size of the subproblem tree. 
5. `early` is the variable specifying whether we use early prediction or not. If `early = 0` then we don't use early prediction strategy, if `early = l` then we perform the early prediction at level `l`.

**Early Prediction**

We can do the early prediction by the following command:

```{r}
dcsvm.model = dcSVM(x = as.matrix(svmguide1[,-1]), y = svmguide1[,1], 
                    k = 10, max.levels = 1, 
                    early = 1, gamma = 2, cost = 32, tolerance = 1e-2, m = 800, 
                    valid.x = svmguide1.t[,-1], valid.y = svmguide1.t[,1])
dcsvm.model$valid.score
dcsvm.model$time$total.time
```

It is faster because we can stop at a level, and don't need to train SVMs for data of larger size.

**Exact Prediction**

To make the model more accurate, we can also perform the exact training by:

```{r}
dcsvm.model = dcSVM(x = as.matrix(svmguide1[,-1]), y = svmguide1[,1], 
                    k = 10, max.levels = 1, 
                    early = 1, gamma = 2, cost = 32, tolerance = 1e-2, m = 800, 
                    valid.x = svmguide1.t[,-1], valid.y = svmguide1.t[,1])
dcsvm.model$valid.score
dcsvm.model$time$total.time
```

This is more accurate but the time is longer. It a balance between the accuracy and the time complexity.

**Sparse Input**

First we download the following dataset.

```{r}
local.file.name = tempfile()
download.file("http://www.sfu.ca/~hetongh/data/ijcnn1.dcsvm.RData",local.file.name)
load(local.file.name)
ijcnn1.t = ijcnn1[[2]]
ijcnn1 = ijcnn1[[1]]
```

These two objects are sparse matrix of class `Matrix::dgCMatrix`. `dcSVM` also supports using sparse matrix as input. 


```{r}
dcsvm.model = dcSVM(x = ijcnn1[,-1], y = ijcnn1[,1], k = 10, max.levels = 1, seed = 1024,
                    early = 1, gamma = 2, cost = 32, tolerance = 1e-2, m = 5000, scale = FALSE,
                    valid.x = ijcnn1.t[,-1], valid.y = ijcnn1.t[,1])
dcsvm.model$valid.score
dcsvm.model$time$total.time
```

However, one should be careful when dealing with sparse matrix, especially when you want to scale the data. After you have scale the data, it could lose the sparsity and make the training process slower.

```{r}
dcsvm.model = dcSVM(x = ijcnn1[,-1], y = ijcnn1[,1], k = 10, max.levels = 1, seed = 1024,
                    early = 1, gamma = 2, cost = 32, tolerance = 1e-2, m = 5000, scale = TRUE,
                    valid.x = ijcnn1.t[,-1], valid.y = ijcnn1.t[,1])
dcsvm.model$valid.score
dcsvm.model$time$total.time
```


The Mixture of SVM Experts with a Gater Function
=======================================================

#### Algorithm

The algorithm can be described in the following iterative framework:

**Training**

1. Split the data into $M$ random subsets of nearly $N/M$.
2. Train $M$ svm experts seperately on each subset.
3. Each svm expert predict on the entire training data set.
4. Train a neural network model minimizing the Eq. (7) in the original paper.
5. Reconstruct subsets: Ffor each data point
    - sort the experts in descending order according to the weight from the neural network
    - assign the example to the first expert in the list which has less than (N/M+c) data points.
6. Goto step 2 until a termination criterion is fulfilled.

**Test**

1. Make prediction on the test data by the $M$ experts.
2. Make prediction on the experts' output by the trained neural network model.

#### Basic usage

```{r}
gaterSVM.model = gaterSVM(x = svmguide1[,-1], y = svmguide1[,1], hidden = 10, seed = 0,
                          m = 10, max.iter = 3, learningrate = 0.01, threshold = 1, stepmax = 1000,
                          valid.x = svmguide1.t[,-1], valid.y = svmguide1.t[,1], verbose = TRUE)
gaterSVM.model$valid.score
```

The parameters can be categorized into the following groups: 

1. `x` and `y` are the feature matric and target vector of the training data. 
2. `valid.x` and `valid.y` are the feature matric and target vector of the validation data.
3. `seed` is controlling the random seed to make the result reproducible. `verbose` is controlling the content of the output.
4. `m`, `max.iter` controls the iteration of "experts-gater" process.
5. `hidden`, `learningrate`, `threshold`, `stepmax` are parameters for the neural network model.





